---
title: "EDA"
output: html_document
author: "Haneef Meeran"
date: "6/24/2019"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    df_print: paged
    highlight: tango
    theme: cosmo
    toc: yes
---

### Data Description

- Country describes the primary variable describes the location of the life expectancy measured

- Region describes the continent related to the country

- Year - Provided for each country from 1800 to 2015

- Life - Provided for each country by year

- Population - Data available for each starting decade of each country

- Income - Continous data series for each country by year


###### Income - Gross domestic product per person adjusted for differences in purchasing power (GDP per capita, PPP$ inflation adjusted)
###### Life - The average number of years a newborn child would live if  current mortality pattern were to stay the same
###### population - Population for the given country in the given year


### Set working directory and Import raw data

```{r}
setwd("~/Desktop/EXPL DATA ANALYSIS/")

gapminder <- read.csv("gapminder.csv")

```

### Required Libraries
```{r}

library(dplyr)

library(ggplot2)

library(matrixStats)

library(scales)

```


### Summary of data
```{r summary}
#Summary of  Gapminder 
summary(gapminder)

str(gapminder)
```

###Remove blanks
```{r}
#Number of Unique Countries
length(unique(gapminder$Country))
#Cleaning of population field. Removing commas and reloading as numeric in a different dataset without NAs.
gapminder$population1 <- as.numeric(gsub(",","",gapminder$population))

gapminder1 <- gapminder %>% filter(!is.na(gapminder$population1))

```


### Questions What is correlation between Income and Population by age groups
### Analysis of Income and Population
```{r income study}

hist(gapminder$income)

summary(unique(gapminder$income))

weightedMedian(gapminder$income, w = gapminder$region, na.rm = TRUE)

weightedMean(gapminder$income, w = gapminder$region, na.rm = TRUE)

d <- ggplot(gapminder1, aes(x = population1, y = income))
d + geom_point()+geom_smooth(method = "lm") + scale_x_continuous(labels = scales::comma) 

```


```{r Cluster Analysis, warning = FALSE, echo=TRUE}

gap_cluster <- select(gapminder1, income, life)


```
Looking at the histogram, a high frequency of income in the range below 50000.

The Histogram shows a very high life expectancy starting at 20, the highest being 20 to 40 and an average income of about 2000 as we observe towards 80.

We can observe that as the population increases, the income hardly increases. If the population is lower, the income increases exponentially. 

###Question Compare population between India and China

##### compare the top most populations of the world, Population between India and China
```{r population between india and china}
# Creating subset of India Country
population_india <- gapminder1 %>% filter(Country=='India')
# Creating subset of India Country
population_china <- gapminder1 %>% filter(Country=='China')

ggplot(data= population_india, mapping = aes(y=Year, x=population1, color="India"))  + geom_point(data = gapminder1, aes(color="Others")) + geom_jitter()

ggplot(data= population_china, mapping = aes(y=Year, x=population1, color="China")) + geom_point(data = gapminder1, aes(color="Others")) + geom_jitter()

```
The population in India and China has risen exponentially after 1950.

###Questions Let find the highest GDP per Capita

##### Data Analysis of Top Countries with High GDP Per Capita from 1990
``` {r High GDP, warning = FALSE}
gdp <- select(gapminder1, Country, Year, income) %>% filter(Year > 1990) %>% group_by(Country) %>% summarise(avg = mean(income)) %>% arrange(desc(avg))

gdp <- gdp[1:8,]

gdp_top <- filter(gapminder1, Country == gdp$Country) %>% filter(Year > 1990)

ggplot(gdp_top, aes(Country, income, color = Country)) +
  geom_boxplot() + labs(title = "Top Countries with High GDP Per Capita", y = "Year", x = "Country_Income")
```

###What is the Life expectancy by continent?
##### Data Analysis of Life expectany by continent
```{r Life-expectancy}

hist(gapminder$life)

summary(gapminder$life)

gapminder_2015 %>%
 ggplot(aes(x = life, fill = region)) + facet_wrap(~Year) + # aes = aesthetics
 geom_density(alpha = 0.7) + 
 
  ggtitle('Life Expectancy by continent') +
 theme(legend.title = element_text(color = 'black',size = 14, face = 'bold'),
 legend.background = element_rect(fill = 'gray90', size = 0.5, linetype = 'dashed')) +
 labs(x='Life Expected (Years)', y='Density')
```
The Histogram shows a very high life expectancy starting at 20, the highest being 20 to 40 and an average income of about 2000 as we observe towards 80.


#### Summary
**Income**<br />
Income for some of the regions are higher than compared to other regions. From the plots above (population vs income), we can also find that as the population increases, the income hardly increases for most, while is linear for some. However, if the population is lower always, the income increases exponentially.


**GDP Per Capita**<br />
We also analysed the Top Countries with the highest GDP per capita from the year 1990, 


**Life Expectancy**<br />
We find that life expectancy in the given dataset is higher in the range of 25-35 years. The mean of Life Expectancy in 2015 for Sub-Saharan Africa region is lower than all other regions, while Europe & Central Asia is highest. 
**There is a relationship between Life Exepectancy and income. 
As the life expectancy grows above 60, the income levels rises higher.

When magnified the analysis to American region, the life Expectancy is almost 70 to 75. 

Though the average life expectancy of America improve from below average to above average from 1800 to 2015, the life expectancy of United States has been highest compared to others. 

